STJ2012 analysis

### 2016-07-29 scripts/STJ2012_analysis.Rmd
move .Rmd to scripts directory

```{r data_import}
# Import OTU table
OTU <- otu_table(read.table("data/otu_table.tsv", header=T, check.names=F, row.names=1,
                            skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
# Import taxonomy
TAX <- tax_table(as.matrix(read.table("data/tax_assignments.txt", sep="\t", row.names=2,
                                      col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
# Import sample data
SAM <- sample_data(read.delim("data/mapping_file.txt", sep="\t", header=T, row.names=1))
# Build phyloseq object
phy <- phyloseq(OTU, TAX, SAM)
phy
```

@@

#### Filter samples by minimum count
```{r filter_samples}
# Set threshold number of reads
sn <- 100
# Remove samples with fewer reads than threshold
phy.f <- prune_samples(sample_sums(phy.f)>=sn, phy.f)

@@

@@

## Clade composition in all samples
```{r cladecomp, fig.height=6, fig.width=6}
samdat <- data.frame(sample_data(phy.f.p))
samdat$Genus <- factor(samdat$Genus, levels=rev(levels(samdat$Genus)))
samdat$Species <- factor(samdat$Species, levels=rev(levels(samdat$Species)))
samdat$Shore <- factor(samdat$Shore, levels=rev(levels(samdat$Shore)))
samdat <- samdat[with(samdat, order(Genus, Species, Shore)), ]
typerelabund <- as.matrix(otu_table(phy.f.p)[order(data.frame(tax_table(phy.f.p))$Subtype), rownames(samdat)])
shorebreaks <- c(as.character(samdat$Shore), "X")==c("X", as.character(samdat$Shore))
shorebreaks <- which(shorebreaks==F) - 1
spbreaks <- c(which(duplicated(samdat$Species)==F) - 1, nrow(samdat))
# Make Barplot
par(mar=c(2, 1.5, 2, 5), lwd=0.1, cex=0.7, mfrow=c(1, 1))
barplot(typerelabund, horiz=T, space=0, axes=F,axisnames=F, yaxs="i",
        col=rainbow(75))#,
        #col=taxcolors[data.frame(tax_table(phy.f.p))[order(data.frame(tax_table(phy.f.p))$Subtype), ]$Clade])
rect(0, 0, par("usr")[2], par("usr")[4], lwd=1, xpd=T)
axis(side=1, at=seq(0, 1, 0.1), line=0, tck=-0.025, mgp=c(0,0.25,0), cex.axis=0.7)
mtext(side=1, "Relative abundance", cex=0.7, line=1)
legend(x=par("usr")[2]/2, y=par("usr")[4], xjust=0.5, yjust=0.25, horiz=T, bty="n", xpd=T, cex=0.7,
       legend=c("A", "B", "C", "D", "F"), fill=taxcolors, x.intersp=0.5)
legend(x=par("usr")[2]-0.03, y=par("usr")[4], xjust=0, yjust=0.1, bty="n", xpd=T, cex=0.6, pt.cex=0.6,
       legend=c("N Shore", "S Shore"), fill=c(0, "gray40"), y.intersp=0.7, x.intersp=0.3)
# Add grouping bars for Shore
for (i in 1:length(shorebreaks)) {
  lines(c(0, 1), c(shorebreaks[i], shorebreaks[i]), lty=2, lwd=0.25)
  rect(1.01, shorebreaks[i], 1.04, shorebreaks[i+1], col=rep(c("gray40", 0), 50)[i], lwd=0.25, xpd=T)
}
# Add lines to separate species and species names
for (i in 1:length(spbreaks)) {
  lines(c(0, 1.07), c(spbreaks[i], spbreaks[i]), xpd=T, type="l", lwd=0.4)
  text(1.03, (spbreaks[i] + spbreaks[i+1]) / 2, xpd=T, pos=4, cex=0.6,
       labels=paste(samdat$Genus[which(duplicated(samdat$Species)==F)][i], "\n",
                    samdat$Species[which(duplicated(samdat$Species)==F)][i], sep=""))
}
```

# Statistical comparison of beta diversity defined by multivariate dispersion of B-C dissimilarity
#   Calculate distances to group centroids in full-dimensional principal coordinate space
#   Implemented by vegan betadisper(), based on Anderson (2006).
# Compartments/species:
```{r }
sambd <- betadisper(d=vegdist(t(otu_table(phy.f.t)), method="bray"), 
                    group=data.frame(sample_data(phy.f.t))$Species, type="centroid")
anova(sambd)
boxplot(sambd)
sambdmeans <- aggregate(sambd$distances, by=list(group=sambd$group), FUN=mean)
sambdses <- aggregate(sambd$distances, by=list(group=sambd$group), 
                      FUN=function(x) sd(x)/sqrt(length(x)))
sampt <- permutest(sambd, pairwise = T)
sampt2 <- TukeyHSD(sambd)
library(multcompView)
saml <- multcompLetters(sampt$pairwise$permuted)#, Letters=c("e","a", "f","d","c","b"))
saml2 <- multcompLetters(sampt2$group[, 4])

>>>>

```{r data_import, cache=T}
# Import OTU table
OTU <- otu_table(read.table("data/otu_table.tsv", header=T, check.names=F, row.names=1,
                            skip=1, sep="\t", comment.char=""), taxa_are_rows=T)
# Import taxonomy
TAX <- tax_table(as.matrix(read.table("data/tax_assignments.txt", sep="\t", row.names=2,
                                      col.names=c("Clade", "OTU.ID", "Taxonomy", "Eval", "Subtype"))))
# Import sample data
SAM <- sample_data(read.delim("data/mapping_file.txt", sep="\t", header=T, row.names=1))
# Build phyloseq object
phy <- phyloseq(OTU, TAX, SAM)
phy
```

@@

#### Filter samples by minimum count
```{r filter_samples}
# Set threshold number of reads
sn <- 10
# Remove samples with fewer reads than threshold
phy.f <- prune_samples(sample_sums(phy.f)>=sn, phy.f)

@@

#### Distribution of OTU counts and reads per sample post-filtering
```{r phy.f_histograms, fig.height=3, fig.width=8}
taxhist <- hist(log10(taxa_sums(phy.f)), plot=F)
samhist <- hist(log10(sample_sums(phy.f)), plot=F)

* Range of OTU counts: `r range(taxa_sums(phy.f))`  
* Number of singletons:  `r length(taxa_sums(phy.f)[taxa_sums(phy.f) <= 1])`  
* Range of reads per sample: `r range(sample_sums(phy.f))`  
* Mean (±SD) reads per sample (arithmetic): `r mean(sample_sums(phy.f))` ± `r sd(sample_sums(phy.f))`  
* Mean (±SD) reads per sample (geometric): `r exp(mean(log(sample_sums(phy.f))))` ± `r exp(sd(log(sample_sums(phy.f))))`

```{r reads_per_sample_filtered, fig.height=2, fig.width=8}
plot_bar(phy.f, x="Sample", y="Abundance")
```

@@

## Community composition in all samples
Here the composition of each sample is plotted as a horizontal bar, sorted by species and shore. Each segment of the bar represents an individual OTU. In the lefthand plot, all OTUs are colored by clade, in the right OTU, all OTUs are given a unique color.
```{r cladecomp, fig.height=6, fig.width=10}
# Define function to plot Symbiodinium community composition
composition <- function(phy, col, legend=T) {
  samdat <- data.frame(sample_data(phy))
  samdat$Genus <- factor(samdat$Genus, levels=rev(levels(samdat$Genus)))
  samdat$Species <- factor(samdat$Species, levels=rev(levels(samdat$Species)))
  samdat$Shore <- factor(samdat$Shore, levels=rev(levels(samdat$Shore)))
  samdat <- samdat[with(samdat, order(Genus, Species, Shore)), ]
  typerelabund <- as.matrix(otu_table(phy)[order(data.frame(tax_table(phy))$Subtype), rownames(samdat)])
  shorebreaks <- c(as.character(samdat$Shore), "X")==c("X", as.character(samdat$Shore))
  shorebreaks <- which(shorebreaks==F) - 1
  spbreaks <- c(which(duplicated(samdat$Species)==F) - 1, nrow(samdat))
  # Make Barplot
  barplot(typerelabund, horiz=T, space=0, axes=F,axisnames=F, yaxs="i", col=col)
  rect(0, 0, par("usr")[2], par("usr")[4], lwd=1, xpd=T)
  axis(side=1, at=seq(0, 1, 0.1), line=0, tck=-0.025, mgp=c(0,0.25,0), cex.axis=0.7)
  mtext(side=1, "Relative abundance", cex=0.7, line=1)
  # Add legend
  if (legend==T) {
    legend(x=par("usr")[2]/2, y=par("usr")[4], xjust=0.5, yjust=0.25, horiz=T, bty="n", xpd=T, cex=0.7, legend=c("A", "B", "C", "D", "F"), fill=taxcolors, x.intersp=0.5)
    legend(x=par("usr")[2]-0.03, y=par("usr")[4], xjust=0, yjust=0.1, bty="n", xpd=T, cex=0.6, pt.cex=0.6, legend=c("N Shore", "S Shore"), fill=c(0, "gray40"), y.intersp=0.7, x.intersp=0.3)
  }
  # Add grouping bars for Shore
  for (i in 1:length(shorebreaks)) {
    lines(c(0, 1), c(shorebreaks[i], shorebreaks[i]), lty=2, lwd=0.25)
    rect(1.01, shorebreaks[i], 1.04, shorebreaks[i+1], col=rep(c("gray40", 0), 50)[i], lwd=0.25, xpd=T)
  }
  # Add lines to separate species and species names
  for (i in 1:length(spbreaks)) {
    lines(c(0, 1.07), c(spbreaks[i], spbreaks[i]), xpd=T, type="l", lwd=0.4)
    text(1.03, (spbreaks[i] + spbreaks[i+1]) / 2, xpd=T, pos=4, cex=0.6,
        labels=paste(samdat$Genus[which(duplicated(samdat$Species)==F)][i], "\n",
                    samdat$Species[which(duplicated(samdat$Species)==F)][i], sep=""))
  }
}

par(mfrow=c(1,2), mar=c(2, 1.5, 2, 5), lwd=0.1, cex=0.7)
composition(phy.f.p, col=taxcolors[data.frame(tax_table(phy.f.p))[order(data.frame(tax_table(phy.f.p))$Subtype), ]$Clade], legend=T)
composition(phy.f.p, col=rainbow(ntaxa(phy.f.p)), legend=F)

# Statistical comparison of beta diversity defined by multivariate dispersion of B-C dissimilarity
#   Calculate distances to group centroids in full-dimensional principal coordinate space
#   Implemented by vegan betadisper(), based on Anderson (2006).
# Compartments/species:
```{r beta, cache=T}
sambd <- betadisper(d=vegdist(t(otu_table(phy.f.t)), method="bray"), 
                    group=data.frame(sample_data(phy.f.t))$Species, type="centroid")
anova(sambd)
boxplot(sambd)
sambdmeans <- aggregate(sambd$distances, by=list(group=sambd$group), FUN=mean)
sambdses <- aggregate(sambd$distances, by=list(group=sambd$group), 
                      FUN=function(x) sd(x)/sqrt(length(x)))
sampt <- permutest(sambd, pairwise = T)
sampt2 <- TukeyHSD(sambd)
library(multcompView)
saml <- multcompLetters(sampt$pairwise$permuted)#, Letters=c("e","a", "f","d","c","b"))
saml2 <- multcompLetters(sampt2$group[, 4])
